---
title: "assignment 8"
author: "Yong Qiao"
date: "10/31/2020"
output: pdf_document
---
\newcommand{\E}{\mathbb{E}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\sgn}{\mathrm{sgn}}
\newcommand{\bx}{\textbf{x}}
\newcommand{\bX}{\textbf{X}}
\newcommand{\bz}{\textbf{z}}
\newcommand{\bbeta}{\boldsymbol{\beta}}
\newcommand{\bPsi}{\boldsymbol{\Psi}}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Orstein--Uhlenbeck Process
Consider the Ornstein-Uhlenbeck process
\begin{align*}
  \dd r(t) = \alpha(b - r(t))\,\dd t + \sigma\,\dd W(t),
\end{align*}
where $\alpha > 0$, $\sigma > 0$, and $b$ are constants.

1. Show that for $t>0$ and $\Delta>0$,
  \begin{align*}
    r(t+\Delta)
    =
    e^{-\alpha\Delta} r(t) + b(1-e^{-\alpha\Delta}) + 
    \frac{\sigma}{\sqrt{2\alpha}} \sqrt{1-e^{-2\alpha\Delta}} Z,
  \end{align*}
  where $Z\sim N(0,1)$.

1. Use the transition distribution from the last part
  to implement a random walk construction for the
  process on time interval $[0,T]$.  Your code should take $\alpha$, 
  $\sigma$, $b$, the initial value $r(0)$, $T$, and the time step
  $\Delta$ of the random walk as input arguments. For $r(0)=1$,
  $T=500$, and $\Delta = 1/500$, plot a sample path for each
  combination of the following values,
  \begin{align*}
    \alpha \in \{0.1, 1, 5\},\ \sigma \in \{0.1, 0.2, 0.5\},\
    b\in\{-5, 5\}.
  \end{align*}
  Comment on how the behavior of $r(t)$ depends on $\alpha$ and
  $\sigma$.

1. Use the Euler–Maruyama method (or the Euler method; see Wiki) to approximate
  a simulation from the process. Specifically, partition the time interval
  into a grid with subintervals of equal length $\delta > 0$ for a small $\delta$;
  approximate $r(t + \delta)$ by a normal random variable with mean
  $r(t) + \alpha(b - r(t)) \delta$ and standard deviation $\sigma \delta$.
  Write a function to implement this approximation with $\delta$ as one of the
  arguments. For $\delta \in \{1, 0.5, 0.1, 0.01\}$, generate a sample of size 1000
  for $r(1)$. Plot the kernel densities against the true density.
  
### Solution

1. The general case of Gaussian Markob process is the following
for the model $$\dd r(t)=[g(t)+h(t)r(t)]\dd t+\sigma (t)\dd W (t)$$
if $H(t)=\int^{t}_{0}h(s)ds$
so we have the general solution is 
$$r(t)=e^{H(t)}r(0)+\int^{t}_{0}e^{H(t)-H(s)}g(s)\dd s+\int^{t}_{0}e^{H(t)-H(s)}\sigma (s)\dd W(s)$$
in this case we have
$$r(t)=e^{\alpha t}r(0)+\int^{t}_{0}e^{\alpha (t-s)}b\dd s+\int^{t}_{0}e^{\alpha(t-s)}\sigma (s)\dd W(s)$$
$$r(t+\Delta)=e^{\alpha (t+\Delta)}r(0)+\int^{t+\Delta}_{0}e^{\alpha (t+\Delta-s)}b\dd s+\int^{t+\Delta}_{0}e^{\alpha(t+\Delta-s)}\sigma (s)\dd W(s)$$
so we have the following
$$r(t+\Delta) - e^{-\alpha \Delta} r(t) = b(1- e^{-\alpha \Delta}) + \sigma \int^{t+\Delta}_t e^{-\alpha(t+\Delta-s)}dW(s)$$
because of $W(s+\Delta s) - W(s) \sim N$ with mean 0

$$E[(\int^{t+\Delta}_t e^{-\alpha (t+\Delta-s)}dW(s))^2] = E[\int^{t+\Delta}_t e^{-2\alpha (t+\Delta-s)}ds] = \frac{(1-e^{-2 \alpha \Delta})}{2\alpha}$$
$$\sigma \int^{t+\Delta}_t e^{-\alpha(t+\Delta-s)}dW(s) \sim N(0, \frac{\sigma^2}{2\alpha}(1-e^{-2 \alpha \Delta}))$$

\begin{align*}
    r(t+\Delta)
    =
    e^{-\alpha\Delta} r(t) + b(1-e^{-\alpha\Delta}) + 
    \frac{\sigma}{\sqrt{2\alpha}} \sqrt{1-e^{-2\alpha\Delta}} Z,
  \end{align*}
  where $Z\sim N(0,1)$.\


2. 
```{r}
ou <- function(alpha, sigma, b, r0, t.max, delta) {
  n <- t.max/delta
  r <- rep(r0, (n + 1))
  for(i in 1:n) {   
    r[i+1] <- exp(-alpha * delta) * r[i] + b * (1 - exp(-alpha * delta)) + 
      sigma * sqrt((1 - exp(-2 * alpha * delta)) / (2 * alpha))  * rnorm(1)
  }
  return(r)
}
a <- c(0.1, 1, 5)
s <- c(0.1, 0.2, 0.5)
b <- c(-5, 5)
r <- matrix(NA, 18, 500^2+1) 
  for (i in 1:2) {
  for (j in 1:3) {
    for (k in 1:3) {
      r[9 * (i - 1) + 3 * (j - 1) + k, ] <- 
        ou(a[j], s[k], b[i], 1, 500, 1/500)
    }    
  }  
}
r.new <- r[, -1]
```

we can plot as the following
```{r}
plot(r.new[1, ],type='l',ylim = c(-9, 2), ylab='', col=2,main='b =  5',xlim=c(0,300000))
lines(r.new[2, ], type='l', col=3)
lines(r.new[3, ], type='l', col=4)
lines(r.new[4, ], type='l', col=5)
lines(r.new[5, ], type='l', col=6)
lines(r.new[6, ], type='l', col=7)
lines(r.new[7, ], type='l', col=8)
lines(r.new[8, ], type='l', col=9)
lines(r.new[9, ], type='l', col=10)
legend("right", legend=c('0.1, 0.1', '0.1, 0.2', '0.1, 0.5','1, 0.1', '1, 0.2', '1, 0.5','5, 0.1', '5, 0.2', '5, 0.5'),lty=c(1,1), col=c(2, 3, 4,5,6,7,8,9,10), title="(alpha , sigma)")
```
```{r}
plot(r.new[10, ],type='l',ylim = c(0, 8), ylab='', col=2,main='b =  -5',xlim=c(0,300000))
lines(r.new[11, ], type='l', col=3)
lines(r.new[12, ], type='l', col=4)
lines(r.new[13, ], type='l', col=5)
lines(r.new[14, ], type='l', col=6)
lines(r.new[15, ], type='l', col=7)
lines(r.new[16, ], type='l', col=8)
lines(r.new[17, ], type='l', col=9)
lines(r.new[18, ], type='l', col=10)
legend("right", legend=c('0.1, 0.1', '0.1, 0.2', '0.1, 0.5','1, 0.1', '1, 0.2', '1, 0.5','5, 0.1', '5, 0.2', '5, 0.5'),lty=c(1,1), col=c(2, 3, 4,5,6,7,8,9,10), title="(alpha , sigma)")
```
From the picture we can see that as our alpha increase the process more stable
From the picture we can see that as our sigma increase the process less stable


3. 
```{r}
euler <- function(alpha, sigma, b, r0, delta) {   
  n <- 1/delta
  r <- matrix(r0, 1000, n + 1) 
  for (i in 1:1000) {
    for (j in 1:n) {
      r[i,j+1]<-r[i,j]+alpha*(b-r[i,j])*delta+sigma*delta*rnorm(1)
      }
  }
  return(r)
}
a <- 0.1
s <- 0.1
b <- 5
init <- 8
delta <- c(1, 0.5, 0.1, 0.01)
mu <- exp(-a * delta[1]) * init + b * (1 - exp(-a * delta[1]))
sig <- s * sqrt((1 - exp(-2 * a * delta[1])) / (2 * a))
true <- function (x) {
  exp(-((x-mu)^2) /(2*sig^2)) / sqrt(2*pi*sig^2)
}
delta_is_1 <- euler(a, s, b, init, delta[1])[, 2]
delta_is_0.5 <- euler(a, s, b, init, delta[2])[, 3]
delta_is_0.1 <- euler(a, s, b, init, delta[3])[, 11]
delta_is_0.01 <- euler(a, s, b, init, delta[4])[, 101]

```

```{r}
plot(density(delta_is_1), col = 1, lwd = 1,ylim=c(0,5))
curve(true, add = T, col = 2)
legend('right', legend = c("simulated","true"),lty = c(1, 1), lwd = c(1, 1), col = 1:2)
```
```{r}
plot(density(delta_is_0.5), col = 1, lwd = 1,ylim=c(0,10))
curve(true, add = T, col = 2)
legend('right', legend = c("simulated","true"),lty = c(1, 1), lwd = c(1, 1), col = 1:2)
```
```{r}
plot(density(delta_is_0.1), col = 1, lwd = 1,ylim=c(0,15))
curve(true, add = T, col = 2)
legend('right', legend = c("simulated","true"),lty = c(1, 1), lwd = c(1, 1), col = 1:2)
```
```{r}
plot(density(delta_is_0.01), col = 1, lwd = 1,ylim=c(0,50))
curve(true, add = T, col = 2)
legend('right', legend = c("simulated","true"),lty = c(1, 1), lwd = c(1, 1), col = 1:2)
```

### Poisson Process

Let $\lambda(t) = \sqrt{t} + e^{-t} \sin(2 \pi t)$
be the intensity function of Poisson process over $t \in [0, 5]$.
Let $N(t)$ be the number of events by time $t$.

1. What is the distribution of $N(5)$ and its parameter(s)?
  Use Mathematica or Maple for integration if needed.
  
1. Write a function to simulate from this Poisson process.

1. Generate events from this Poisson process 1,000 times. Pool all
  the event points together as a sample and plot their kernel density. 
  Overlay $\lambda(t) / \int_0^5 \lambda(s) \dd s$ with the kernel density.
  

### Solution

$N(5) \sim Pisson(Z)$

$Z =\int^5_0 \lambda(t)dt = \int^5_0 (\sqrt{t} + e^{-t} \sin(2 \pi t))dt$
$N(5) \sim Pisson(Z)$

$$
\begin{aligned}
\int^5_0 e^{-t} \sin(2 \pi t)dt 
  &= \frac{(2\pi)^2}{1+(2\pi)^2}(\frac{1}{2\pi} - \frac{e^{-5}}{2\pi})\\
  &= \frac{(2\pi)(1-e^{-5})}{1+(2\pi)^2}
\end{aligned}
$$
$Z = \frac{2}{3} \times 5^{3/2} + \frac{(2\pi)(1-e^{-5})}{1+(2\pi)^2}$
$N(5) \sim Pisson(Z)=Pisson(\frac{2}{3} \times 5^{3/2} + \frac{(2\pi)(1-e^{-5})}{1+(2\pi)^2})$

```{r}
t.max <- 5 
lambda.0 <- sqrt(t.max) + exp(-t.max)*sin(2*pi*t.max) 
lambda <- function(t) {
  sqrt(t) + exp(-t)*sin(2*pi*t)
} 
y <- c() 
sim_poission_q2 <- function(n) {
  for(i in 1:n) { 
    n <- rpois(1, lambda.0 * 5) # N
    if (n > 0) {
      tau = 5 * sort(runif(n))
      for (i in 1:length(tau)) {
        t <- tau[i]
        if( lambda(t) / lambda.0 > runif(1))
          y = c(y,t)
      }
    }
  }
  return(y)
}  
```

```{r}
summulation_1000  <- sim_poission_q2(1000) 
true_process <- function(t) {
  (sqrt(t) + exp(-t)*sin(2*pi*t)) / (2*5^(1.5)/3 + 2*pi*(1-exp(-5))/(1+4*(pi^2)))
} 
summulation_1000 <- density(summulation_1000)
plot(summulation_1000,col=2,xlim=c(-1,7),ylim=c(0,0.5))
curve(true_process, 0, 5, add = T,col=3)
legend('topright', legend = c("simulated","true"),
       lty = c(1,1), lwd = c(1,1), col=c(2,3),
       cex = 0.9)
```

